---
title: "Topic 07 - Word Embeddings"
author: "Julia Parish"
date: '2022-05-11'
header-includes:
  - \usepackage{float}
  - \floatplacement{figure}{H}
output: 
  pdf_document:
    extra_dependencies: ["float"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.pos='!H')
```

# Word Embedding

This text sentiment analysis was completed as an assignment for the course, Environmental Data Science 231: Text and Sentiment Analysis for Environmental Problems. The data was sourced from: Jeffrey Pennington, Richard Socher, and Christopher D. Manning. 2014. GloVe: Global Vectors for Word Representation.

Original assignment instructions can be found [here](https://maro406.github.io/EDS_231-text-sentiment/topic_7.html)

### Load Libraries

```{r}
if (!require(librarian)){
  install.packages("librarian")
  library(librarian)
}

librarian::shelf(broom,
                 here,
                 irlba,
                 kableExtra,
                 textdata,
                 tidytext,
                 tidyverse,
                 widyr)
  
```

