---
title: "Topic 04 - Sentiment Analysis II"
author: "Julia Parish"
date: '2022-04-26'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Sentiment Analyis II

This text sentiment analysis was completed as an assignment for the course, Environmental Data Science 231: Text and Sentiment Analysis for Environmental Problems. The data was sourced from Twitter. 

Original assignment instructions can be found [here](https://maro406.github.io/EDS_231-text-sentiment/topic_4.html)

### Load Library
```{r}
library(quanteda)
library(quanteda.sentiment)
library(quanteda.textstats)
library(tidyverse)
library(tidytext)
library(lubridate)
library(wordcloud)
library(reshape2)
library(here)
```

## Load IPPC tweet data & create plot of data
```{r ipcc tweet data}
raw_tweets <- read.csv(here("assignments/HW04_SentimentAnalysisII/data/IPCC_tweets_April1-10_sample.csv"))

dat<- raw_tweets[,c(5,7)] # Extract Date and Title fields

tweets <- tibble(text = dat$Title,
                  id = seq(1:length(dat$Title)),
                 date = as.Date(dat$Date,'%m/%d/%y'))


head(tweets$text, n = 10)

#simple plot of tweets per day
tweets %>%
  count(date) %>%
  ggplot(aes(x = date, y = n))+
  geom_line() +
  labs(title = "IPCC Tweets per Day",
       subtitle = "April 01 - April 10, 2022",
       x = "Date",
       y = "Number of Tweets") + 
  theme_minimal()
```


# Questions

## 1. Think about how to further clean a twitter data set. Let’s assume that the mentions of twitter accounts is not useful to us. Remove them from the text field of the tweets tibble.

## 2. Compare the ten most common terms in the tweets per day. Do you notice anything interesting?

## 3. Adjust the wordcloud in the “wordcloud” chunk by coloring the positive and negative words so they are identifiable.

## 4. Let’s say we are interested in the most prominent entities in the Twitter discussion. Which are the top 10 most tagged accounts in the data set. Hint: the “explore_hashtags” chunk is a good starting point.

## 5. The Twitter data download comes with a variable called “Sentiment” that must be calculated by Brandwatch. Use your own method to assign each tweet a polarity score (Positive, Negative, Neutral) and compare your classification to Brandwatch’s (hint: you’ll need to revisit the “raw_tweets” data frame).