---
title: "Lab01_NYTimes"
author: "Julia Parish"
date: "4/6/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Connect to the New York Times API and send a query
```{r}
library(jsonlite) #convert results from API queries into R-friendly formats 
library(tidyverse) 
library(tidytext) #text data management and analysis
library(ggplot2) #plot word frequencies and publication dates
```

We have to decide which New York Times articles we are interested in examining. For this exercise, I chose articles about Deb Haaland, the current US Secretary of the Interior. As a member of the Laguna Pueblo Tribe, Haaland is the first Native American to serve as Cabinet secretary. 

```{r}
# create an object called x with the results of our query ("haaland")
# the from JSON flatten the JSON object, then convert to a data frame
t <- fromJSON("http://api.nytimes.com/svc/search/v2/articlesearch.json?q=haaland&api-key=qUngwwaLSVRerrzAB0TgUdXSHf93H6ea", flatten = TRUE) #the string following "key=" is your API key 

# key=NTKBHbsb6XFEkGymGumAiba7n3uBvs8V

class(t) #what type of object is t?

t <- t %>% 
  data.frame()

#Inspect our data
class(t) #now what is it?
dim(t) # how big is it? 10 articles, 33 variables (fields)

names(t) # what variables are we working with?
# response.docs.abstract where docs is a list, and abstract is a list within docs

#t <- readRDS("nytDat.rds") #in case of API emergency :)
```

```{r}
t$response.docs.snippet[9] # one sentence pulled from article (example is 9th one), $ referring to column. snippet is a NYTimes API term

#assign a snippet to x to use as fodder for stringr functions.  You can follow along using the sentence on the next line.

x <- "Her nomination as secretary of the interior is historic, but as the first Native cabinet member, she would have to strike a delicate balance." 

tolower(x) #make all tokens uniform
str_split(x, ','); str_split(x, 't') # x is currently one string, but can split & tell R where to split it.

str_replace(x, 'historic', 'without precedent') # replace token with another

str_replace(x, ' ', '_') #first one
#how do we replace all of them?
str_replace_all(x, ' ', '_') 

str_detect(x, 't'); str_detect(x, 'tive') ### is pattern in the string? T/F

str_locate(x, 't'); str_locate_all(x, 'as')
```

# parameters for bigger query
```{r}
term <- "Haaland" # Need to use + to string together separate words
begin_date <- "20210120"
end_date <- "20220401"

#construct the query url using API operators
baseurl <- paste0("http://api.nytimes.com/svc/search/v2/articlesearch.json?q=",term,
                  "&begin_date=",begin_date,"&end_date=",end_date,
                  "&facet_filter=true&api-key=","qUngwwaLSVRerrzAB0TgUdXSHf93H6ea", sep="")

#examine our query url
baseurl
```

```{r}
#this code allows for obtaining multiple pages of query results 
 initialQuery <- fromJSON(baseurl)
maxPages <- round((initialQuery$response$meta$hits[1] / 10)-1) 

pages <- list()
for(i in 0:maxPages){
  nytSearch <- fromJSON(paste0(baseurl, "&page=", i), flatten = TRUE) %>% data.frame() 
  message("Retrieving page ", i)
  pages[[i+1]] <- nytSearch 
  Sys.sleep(6) 
}
class(nytSearch)

#need to bind the pages and create a tibble from nytDat
# rowbind pages object

nytDat <- rbind_pages(pages)
```

```{r}
nytDat %>% 
  group_by(response.docs.type_of_material) %>%
  summarize(count=n()) %>%
  mutate(percent = (count / sum(count))*100) %>%
  ggplot() +
  geom_bar(aes(y=percent, x=response.docs.type_of_material, fill=response.docs.type_of_material), stat = "identity") + coord_flip()
```

```{r}
nytDat %>%
  mutate(pubDay=gsub("T.*","",response.docs.pub_date)) %>% # T. is a regular expression, removing anywhere where T.* (time) showed up replaced with blank 
  group_by(pubDay) %>%
  summarise(count=n()) %>%
  filter(count >= 2) %>%
  ggplot() +
  geom_bar(aes(x=reorder(pubDay, count), y=count), stat="identity") + coord_flip()
```

```{r}
names(nytDat)

paragraph <- names(nytDat)[6] #The 6th column, "response.doc.lead_paragraph", is the one we want here.  
tokenized <- nytDat %>%
  unnest_tokens(word, paragraph) #unnest paragraph and convert to collection of single words; one row is one paragraph, this expands dataframe and converts each word in paragraph to an individual row. 

tokenized %>%
  count(word, sort = TRUE) %>%
  filter(n > 5) %>% #illegible with all the words displayed; # using filter to show words used 5x
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(n, word)) +
  geom_col() +
  labs(y = NULL)
```

### remove common words that are not meaningful (a, the, of, etc)
```{r}
data(stop_words) # pulls in data object from data package (tidytext)

tokenized <- tokenized %>%
  anti_join(stop_words) # removes rows with matching word b/w two objects (tokenized & stop words)

tokenized %>%
  count(word, sort = TRUE) %>%
  filter(n > 5) %>% # using filter to show words used 5x
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(n, word)) +
  geom_col() +
  labs(y = NULL)

```

```{r}
#inspect the list of tokens (words)
tokenized$word 

clean_tokens <- str_replace_all(tokenized$word,"land[a-z,A-Z]*","land") #stem tribe words
clean_tokens <- str_remove_all(clean_tokens, "[:digit:]") #remove all numbers. :digit: is a regular expression character

clean_tokens <- str_remove_all(clean_tokens, "washington")
clean_tokens <- gsub("’s", '', clean_tokens) # base R function, removing "'s" possesion

tokenized$clean <- clean_tokens

tokenized %>%
  count(clean, sort = TRUE) %>%
  filter(n > 10) %>% #illegible with all the words displayed
  mutate(clean = reorder(clean, n)) %>%
  ggplot(aes(n, clean)) +
  geom_col() +
  labs(y = NULL)

#remove the empty strings
tib <-subset(tokenized, clean!="")

#reassign
tokenized <- tib

#try again
tokenized %>%
  count(clean, sort = TRUE) %>%
  filter(n > 10) %>% #illegible with all the words displayed
  mutate(clean = reorder(clean, n)) %>%
  ggplot(aes(n, clean)) +
  geom_col() +
  labs(y = NULL)
```


